---
title: "Analysis of Full Data Fit"
author: "Jinyu Luo"
date: "2024-07-19"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(kableExtra)
load("~/Documents/Projects/Practicum_GEE/Outputs/fulldata_fit.RData")
load("~/Documents/Projects/Practicum_GEE/data/realcoefs.RData")
```

```{r}
true_set <- red_ar1$coefficients %>% select(Estimate) %>% 
  rename(truth = Estimate) %>% 
  rownames_to_column(var = "term")
```

### Convergence 
```{r}
calculate_convergence <- function(fits_df){
  temp <- fits_df %>%
    mutate(
      ind_full = map_lgl(ind_mdl_full, ~any(.x$convergence == TRUE)),
      ar1_full = map_lgl(ar1_mdl_full, ~any(.x$convergence == TRUE)),
      exch_full = map_lgl(exch_mdl_full, ~any(.x$convergence == TRUE)),
      ind_red = map_lgl(ind_mdl_red, ~any(.x$convergence == TRUE)), 
      ar1_red = map_lgl(ar1_mdl_red, ~any(.x$convergence == TRUE)), 
      exch_red = map_lgl(exch_mdl_red, ~any(.x$convergence == TRUE))
    ) %>%
    select(sim_id, ind_full:exch_red)
  
  converged <- temp %>% select(-sim_id) %>% colMeans()
  diverged <- temp %>% filter(if_all(ind_full:exch_red, ~ . == FALSE)) 
  if(nrow(diverged) != 0){
    diverged_sim <- diverged %>% pull(sim_id)
    out_data <- fits_df %>% filter(!(sim_id %in% diverged))
  } else{
    out_data <- fits_df
    diverged_sim <- NULL 
  }
  return(list(data = out_data, diverged = diverged_sim)) 
}
```

```{r}
gee_convergence <- calculate_convergence(gee_fits_df)
gee_sim_results <- gee_convergence$data

qls_convergence <- calculate_convergence(qls_fits_df)
qls_sim_results <- qls_convergence$data

divergence <- list(gee = gee_convergence$diverged, qls = qls_convergence$diverged)
if(is.null(divergence$gee) & is.null(divergence$qls)){
  cat("All simulations converged")
  }else{divergence}
```


### Extreme SE Search 
```{r}
extr_se_search <- function(sim_results, se_max) {
  extreme_df <- sim_results %>% 
    mutate(ind_full_se = map_lgl(ind_mdl_full, ~any(.x$std_error > se_max)),
           ar1_full_se = map_lgl(ar1_mdl_full, ~any(.x$std_error > se_max)),
           exch_full_se = map_lgl(exch_mdl_full, ~any(.x$std_error > se_max)),
           ind_red_se = map_lgl(ind_mdl_red, ~any(.x$std_error > se_max)), 
           ar1_red_se = map_lgl(ar1_mdl_red, ~any(.x$std_error > se_max)), 
           exch_red_se = map_lgl(exch_mdl_red, ~any(.x$std_error > se_max))) %>% 
    select(sim_id, ind_full_se:exch_red_se) %>% 
    mutate(across(everything(), ~replace_na(., TRUE))) %>% 
    filter(if_any(-sim_id, ~ . == TRUE | is.na(.))) 
  
  n_extreme <- nrow(extreme_df)
  messages <- character()
  
  if(n_extreme != 0){
    messages <- c(messages, paste("There are", n_extreme, "simulations has extreme values."))
    sim_results <- sim_results %>% filter(!(sim_id %in% extreme_df$sim_id))
    messages <- c(messages, paste(1500 - nrow(sim_results), "simulations with extreme S.E. are removed for the fit."))
    messages <- c(messages, "Removed Simulation ID:")
    messages <- c(messages, paste(unlist(extreme_df$sim_id), collapse = ", "))
  } else {
    messages <- c(messages, "No extreme S.E.")
  }
  
  list(
    filtered_results = sim_results,
    messages = messages, 
    extreme_id = unlist(extreme_df$sim_id)
  )
}
```

 
```{r}
gee_extremSE <- extr_se_search(gee_sim_results, se_max = 5)
cat("GEE\n", gee_extremSE$messages)
```

```{r}
qls_extremSE <- extr_se_search(qls_sim_results, se_max = 5)
cat("QLS\n", qls_extremSE$messages)
```


```{r}
extreme_SEs <- list(gee = gee_extremSE$extreme_id, qls = qls_extremSE$extreme_id)
gee_sim_results <- gee_extremSE$filtered_results
qls_sim_results <- qls_extremSE$filtered_results
```


### Coverage Proability

In summary, coverage probabilities based on GEE approach are all above 90% irrespective of the inclusion of confounding covariates. 
```{r}
calculate_coverage <- function(fit_list, true_set) {
  fit_list %>%
    map_dfr(~.x %>%
              right_join(true_set, by = "term") %>%
              mutate(unadjusted = truth >= lower & truth <= upper,
                     adjusted = truth >= adj_lower & truth <= adj_upper)) %>%
    group_by(term) %>%
    summarize(unadjusted = mean(unadjusted, na.rm = TRUE),
              adjusted = mean(adjusted, na.rm = TRUE))
}
```


```{r}
# GEE 
# (1) Coverage probabilities for models without adjustment by age, male, and BSA
coverage_gee_unadj <- calculate_coverage(gee_sim_results$ind_mdl_red, true_set) %>%
  rbind(calculate_coverage(gee_sim_results$ar1_mdl_red, true_set)) %>%
  rbind(calculate_coverage(gee_sim_results$exch_mdl_red, true_set)) %>%
  na.omit()

# (2) Coverage probabilities for models adjusted by age, male, and BSA
coverage_gee_adj <- calculate_coverage(gee_sim_results$ind_mdl_full, true_set) %>%
  rbind(calculate_coverage(gee_sim_results$ar1_mdl_full, true_set)) %>%
  rbind(calculate_coverage(gee_sim_results$exch_mdl_full, true_set)) %>%
  na.omit()

cbind(coverage_gee_unadj, coverage_gee_adj %>% select(-term)) %>% 
  cbind(corstr = c(rep("Independence", 4), rep("AR1", 4), rep("Exchangeable", 4))) %>% 
  relocate(corstr, .before = term) %>% 
  kable(booktabs = TRUE, digits = 3, 
        col.names = c("Working Correlation", "Term", rep(c("Empirical SE", "SE-DF"), 2))) %>% 
  kable_paper("striped", full_width = F) %>%
  kable_styling(latex_options = "HOLD_position") %>% 
  collapse_rows(columns = 1, valign = "top") %>%
  column_spec(1, bold = T) %>%
  add_header_above(c(" " = 2, "No Confounders" = 2, "With Confounders" = 2))
```

However, coverage probabilities from QLS fit are pretty poor, especially for the coefficient of BAV (lower than 35%). After including confounding covariates, improvements are only seen in the coverage for the intercept term but not BAV. The coverage probabilities for the interaction effects are not plausible as well. The highest coverage probability is achieved only when the working correlation is correctly specified, i.e., AR1. 

```{r}
# QLS
# (1) Coverage probabilities for models without adjustment by age, male, and BSA
coverage_qls_unadj <- calculate_coverage(qls_sim_results$ind_mdl_red, true_set) %>%
  rbind(calculate_coverage(qls_sim_results$ar1_mdl_red, true_set)) %>%
  rbind(calculate_coverage(qls_sim_results$exch_mdl_red, true_set)) %>%
  na.omit()

# (2) Coverage probabilities for models adjusted by age, male, and BSA
coverage_qls_adj <- calculate_coverage(qls_sim_results$ind_mdl_full, true_set) %>%
  rbind(calculate_coverage(qls_sim_results$ar1_mdl_full, true_set)) %>%
  rbind(calculate_coverage(qls_sim_results$exch_mdl_full, true_set)) %>%
  na.omit()

cbind(coverage_qls_unadj, coverage_qls_adj %>% select(-term)) %>% 
  cbind(corstr = c(rep("Independence", 4), rep("AR1", 4), rep("Exchangeable", 4))) %>% 
  relocate(corstr, .before = term) %>% 
  kable(booktabs = TRUE, digits = 3, 
        col.names = c("Working Correlation", "Term", rep(c("Empirical SE", "SE-DF"), 2))) %>% 
  kable_paper("striped", full_width = F) %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  collapse_rows(columns = 1, valign = "top") %>%
  column_spec(1, bold = T) %>%
  add_header_above(c(" " = 2, "No Confounders" = 2, "With Confounders" = 2))
```

#### Coverage Visualization

```{r}
extract_term_data <- function(df, term, mod) {
  output <- df %>%
    select(sim_id, ends_with(mod)) %>%
    pivot_longer(cols = ends_with(mod), names_to = "model", values_to = "data") %>%
    rowwise() %>%
    mutate(data = list(filter(data, term == !!term))) %>%
    unnest(cols = c(data)) %>%
    ungroup() %>%
    mutate(sim_id = as.integer(sim_id), 
           model = case_when(str_detect(model, "ind") ~ "Independence", 
                             str_detect(model, "ar1") ~ "AR1", 
                             TRUE ~ "Exchangeable")) %>% 
    select(sim_id, model, term, estimate, std_error, lower, upper, adj_lower, adj_upper)
  return(output)
}
```

```{r}
bav_red <- extract_term_data(gee_sim_results, term = "bav", mod = "_red") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "bav", mod = "_red") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

visit_red <- extract_term_data(gee_sim_results, term = "visit", mod = "_red") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "visit", mod = "_red") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

bav_visit_red <- extract_term_data(gee_sim_results, term = "bav:visit", mod = "_red") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "bav:visit", mod = "_red") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

red_mdl_df <- rbind(bav_red, visit_red, bav_visit_red)

red_summary <- red_mdl_df %>%
  mutate(term = case_when(term == "bav" ~ "BAV",
                          term == "visit" ~ "Visit",
                          TRUE ~ "BAV:Visit")) %>%
  group_by(term, model, type) %>%
  summarise(mean_estimate = mean(estimate),
            mean_lower = mean(lower),
            mean_upper = mean(upper),
            mean_adj_lower = mean(adj_lower),
            mean_adj_upper = mean(adj_upper),
            .groups = 'drop') %>%
  pivot_longer(cols = c(mean_lower, mean_upper, mean_adj_lower, mean_adj_upper),
               names_to = "ci_type", values_to = "ci_value") %>%
  mutate(ci_type = case_when(
    ci_type == "mean_lower" ~ "Unadjusted Lower",
    ci_type == "mean_upper" ~ "Unadjusted Upper",
    ci_type == "mean_adj_lower" ~ "Adjusted Lower",
    ci_type == "mean_adj_upper" ~ "Adjusted Upper")) %>%
  pivot_wider(names_from = ci_type, values_from = ci_value) %>%
  pivot_longer(cols = c("Unadjusted Lower", "Unadjusted Upper", "Adjusted Lower", "Adjusted Upper"),
               names_to = "ci_type", values_to = "ci_value") %>%
  separate(ci_type, into = c("adjusted", "boundary"), sep = " ") %>%
  pivot_wider(names_from = boundary, values_from = ci_value)
```

```{r}
bav_full <- extract_term_data(gee_sim_results, term = "bav", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "bav", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

visit_full <- extract_term_data(gee_sim_results, term = "visit", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "visit", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

bav_visit_full <- extract_term_data(gee_sim_results, term = "bav:visit", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "bav:visit", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

age_df <- extract_term_data(gee_sim_results, term = "age", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "age", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

male_df <- extract_term_data(gee_sim_results, term = "male", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "male", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

bsa_df <- extract_term_data(gee_sim_results, term = "bsa", mod = "_full") %>% mutate(type = "GEE") %>%
  rbind(extract_term_data(qls_sim_results, term = "bsa", mod = "_full") %>% mutate(type = "QLS")) %>%
  relocate(type, .after = sim_id)

full_mdl_df <- rbind(bav_full, visit_full, bav_visit_full, age_df, male_df, bsa_df)

full_summary <- full_mdl_df %>%
  mutate(term = case_when(term == "bav" ~ "BAV",
                          term == "visit" ~ "Visit",
                          term == "age" ~ "Age",
                          term == "male" ~ "Male",
                          term == "bsa" ~ "BSA",
                          TRUE ~ "BAV:Visit")) %>%
  group_by(term, model, type) %>%
  summarise(mean_estimate = mean(estimate),
            mean_lower = mean(lower),
            mean_upper = mean(upper),
            mean_adj_lower = mean(adj_lower),
            mean_adj_upper = mean(adj_upper),
            .groups = 'drop') %>%
  pivot_longer(cols = c(mean_lower, mean_upper, mean_adj_lower, mean_adj_upper),
               names_to = "ci_type", values_to = "ci_value") %>%
  mutate(ci_type = case_when(
    ci_type == "mean_lower" ~ "Unadjusted Lower",
    ci_type == "mean_upper" ~ "Unadjusted Upper",
    ci_type == "mean_adj_lower" ~ "Adjusted Lower",
    ci_type == "mean_adj_upper" ~ "Adjusted Upper")) %>%
  pivot_wider(names_from = ci_type, values_from = ci_value) %>%
  pivot_longer(cols = c("Unadjusted Lower", "Unadjusted Upper", "Adjusted Lower", "Adjusted Upper"),
               names_to = "ci_type", values_to = "ci_value") %>%
  separate(ci_type, into = c("adjusted", "boundary"), sep = " ") %>%
  pivot_wider(names_from = boundary, values_from = ci_value)
```


```{r}
red_summary %>% 
  left_join(
    true_set %>% mutate(term = case_when(term == "visit" ~ "Visit", 
                                         term == "bav" ~ "BAV", 
                                         term == "bav:visit" ~ "BAV:Visit")), 
    by = "term") %>% 
  mutate(se = ifelse(adjusted == "Adjusted", "SE-DF", "Empirical SE")) %>% 
  rename(corstr = model, 
         model = type) %>% 
  ggplot(aes(x = corstr, y = mean_estimate, color = corstr)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2, 
                position = position_dodge(width = 0.5)) +
  geom_hline(aes(yintercept = truth), linetype = "dashed", color = "black") +
  facet_grid(se ~ model + term, scales = "free_y", 
             axes = "all", axis.labels = "all_x") +
  labs(y = "Mean Estimate",
       x = "",
       color = "Working Correlation", 
       caption = "Without adjusting for confounding covariates.") +
  scale_color_brewer(palette = "Pastel1") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"), 
        panel.grid.major.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        strip.placement = "outside", 
        legend.position = "bottom")
```

## Relative Bias 

In summary, the amount of outliers in relative bias for each term does not depends on the covariate set, but relys on the method and specified working correlations. The GEE method tends to produce more outliers in relative bias, but its mean relative bias is around 0, while the QLS method tends to overestimate coefficients. 

```{r}
cal_rel_bias <- function(df, true_set){
 df %>% 
    map_dfr(~.x %>% select(term, estimate)) %>%
    right_join(true_set, by = "term") %>% 
    group_by(term) %>% 
    mutate(bias = estimate - truth, 
           rel_bias = bias/abs(truth), 
           sim = row_number()) %>% 
   ungroup() %>% na.omit() %>% 
   select(sim, term, rel_bias) %>% 
   pivot_wider(names_from = term, values_from = rel_bias) %>% 
   rename(b0 = `(Intercept)`, 
          b1 = bav, 
          b2 = visit, 
          b3 = `bav:visit`)
}
```

```{r}
red_relbias_df <- cal_rel_bias(gee_sim_results$ind_mdl_red, true_set) %>% 
  cbind(method = "GEE") %>% cbind(corstr = "Independence") %>% 
  rbind(cal_rel_bias(gee_sim_results$ar1_mdl_red, true_set) %>% 
          cbind(method = "GEE") %>% cbind(corstr = "AR1")) %>% 
  rbind(cal_rel_bias(gee_sim_results$exch_mdl_red, true_set) %>% 
          cbind(method = "GEE") %>% cbind(corstr = "Exchangeable")) %>% 
  rbind(cal_rel_bias(qls_sim_results$ind_mdl_red, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "Independence")) %>% 
  rbind(cal_rel_bias(qls_sim_results$ar1_mdl_red, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "AR1")) %>% 
  rbind(cal_rel_bias(qls_sim_results$exch_mdl_red, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "Exchangeable")) %>% 
  cbind(cov_set = "No")
full_relbias_df <- cal_rel_bias(gee_sim_results$ind_mdl_full, true_set) %>% 
  cbind(method = "GEE") %>% cbind(corstr = "Independence") %>% 
  rbind(cal_rel_bias(gee_sim_results$ar1_mdl_full, true_set) %>% 
          cbind(method = "GEE") %>% cbind(corstr = "AR1")) %>% 
  rbind(cal_rel_bias(gee_sim_results$exch_mdl_full, true_set) %>% 
          cbind(method = "GEE") %>% cbind(corstr = "Exchangeable")) %>% 
  rbind(cal_rel_bias(qls_sim_results$ind_mdl_full, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "Independence")) %>% 
  rbind(cal_rel_bias(qls_sim_results$ar1_mdl_full, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "AR1")) %>% 
  rbind(cal_rel_bias(qls_sim_results$exch_mdl_full, true_set) %>% 
          cbind(method = "QLS") %>% cbind(corstr = "Exchangeable")) %>% 
  cbind(cov_set = "Yes")

relbias_df <- rbind(red_relbias_df, full_relbias_df) %>% 
  relocate(c(method, corstr, cov_set), .before = b0)
```

### BAV 

The estimates for the effect of BAV are biased using QLS approach, but many outliers produced within GEE method. 

```{r}
relbias_df %>% select(-c(b0, b2, b3)) %>% 
  filter(b1 > -110 & b1 < 100) %>% 
  ggplot(aes(x = corstr, y = b1, color = corstr)) +
  geom_boxplot()+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  facet_grid(cov_set ~ method, scales = "free_y", 
             axes = "all", axis.labels = "all_x") +
  labs(y =expression(beta[1] ~ "(BAV)"),
       x = "",
       title = "Relative Bias for the effect of BAV",
       color = "Working Correlation", 
       caption = "No = not include confounding covariates,
       Yes = include confounding covariates") +
  scale_color_brewer(palette = "Pastel1") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"), 
        panel.grid.major.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        strip.placement = "outside", 
        legend.position = "bottom")
```

Both GEE and QLS method produced the greatest number of outliers when the working correlation is specified to independence. The number of outliers does not depend on the inclusion of the confounding covariate set. 

```{r}
bav_outliers_df <- relbias_df %>%
  select(-c(b0, b2, b3)) %>%
  group_by(corstr, cov_set, method) %>%
  summarise(Q1 = quantile(b1, 0.25), Q3 = quantile(b1, 0.75)) %>%
  mutate(IQR = Q3 - Q1,
         lower_bound = Q1 - 1.5 * IQR,
         upper_bound = Q3 + 1.5 * IQR)

bav_outliers <- relbias_df %>%
  select(-c(b0, b2, b3)) %>%
  left_join(bav_outliers_df, by = c("corstr", "cov_set", "method")) %>%
  filter(b1 < lower_bound | b1 > upper_bound) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)

bav_outliers %>% 
  group_by(method, corstr, cov_set) %>% 
  summarise(n = n()) %>% 
  arrange(n) %>% 
  kable(booktabs = TRUE, digits = 3, 
        col.names = c("Method", "Working Correlation", 
                      "Confounders", "No.Outliers"), 
        caption = "Outliers in Relative Bias for the effect of BAV") %>% 
  kable_paper("striped", full_width = F) %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  collapse_rows(columns = 1:2, valign = "top") 
```


### Visit

QLS still tends to give biased estimates for the effect of Time, while the estimates from GEE method are very robust with 0 relative bias. 

```{r}
relbias_df %>% select(-c(b0, b1, b3)) %>% 
  filter(b2 > -3 & b2 < 3) %>% 
  ggplot(aes(x = corstr, y = b2, color = corstr)) +
  geom_boxplot()+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  facet_grid(cov_set ~ method, scales = "free_y", 
             axes = "all", axis.labels = "all_x") +
  labs(y =expression(beta[2] ~ "Time"),
       x = "",
       title = "Relative Bias for the effect of Time",
       color = "Working Correlation", 
       caption = "No = not include confounding covariates,
       Yes = include confounding covariates") +
  scale_color_brewer(palette = "Pastel1") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"), 
        panel.grid.major.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        strip.placement = "outside", 
        legend.position = "bottom")
```


```{r}
visit_outliers_df <- relbias_df %>%
  select(-c(b0, b1, b3)) %>%
  group_by(corstr, cov_set, method) %>%
  summarise(Q1 = quantile(b2, 0.25), Q3 = quantile(b2, 0.75)) %>%
  mutate(IQR = Q3 - Q1,
         lower_bound = Q1 - 1.5 * IQR,
         upper_bound = Q3 + 1.5 * IQR)

visit_outliers <- relbias_df %>%
  select(-c(b0, b1, b3)) %>%
  left_join(visit_outliers_df, by = c("corstr", "cov_set", "method")) %>%
  filter(b2 < lower_bound | b2 > upper_bound) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)


visit_outliers %>% 
  group_by(method, corstr, cov_set) %>% 
  summarise(n = n()) %>% 
  arrange(n) %>% 
  kable(booktabs = TRUE, digits = 3, 
        col.names = c("Method", "Working Correlation", 
                      "Confounders", "No.Outliers"), 
        caption = "Outliers in Relative Bias for the effect of Time") %>% 
  kable_paper("striped", full_width = F) %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  collapse_rows(columns = 1:2, valign = "top") 
```

### Interaction effect 
```{r}
relbias_df %>% select(-c(b0, b1, b2)) %>% 
  filter(b3 > -3 & b3 < 3) %>% 
  ggplot(aes(x = corstr, y = b3, color = corstr)) +
  geom_boxplot()+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  facet_grid(cov_set ~ method, scales = "free_y", 
             axes = "all", axis.labels = "all_x") +
  labs(y =expression(beta[3] ~ "(BAV" %*% "Time)"),
       x = "",
       title = "Relative Bias for the interaction effect between BAV and Time",
       color = "Working Correlation", 
       caption = "No = not include confounding covariates,
       Yes = include confounding covariates") +
  scale_color_brewer(palette = "Pastel1") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"), 
        panel.grid.major.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        strip.placement = "outside", 
        legend.position = "bottom")
```


```{r}
interaction_outliers_df <- relbias_df %>%
  select(-c(b0, b1, b2)) %>%
  group_by(corstr, cov_set, method) %>%
  summarise(Q1 = quantile(b3, 0.25), Q3 = quantile(b3, 0.75)) %>%
  mutate(IQR = Q3 - Q1,
         lower_bound = Q1 - 1.5 * IQR,
         upper_bound = Q3 + 1.5 * IQR)

interaction_outliers <- relbias_df %>%
  select(-c(b0, b1, b2)) %>%
  left_join(interaction_outliers_df, by = c("corstr", "cov_set", "method")) %>%
  filter(b3 < lower_bound | b3 > upper_bound) %>%
  select(-Q1, -Q3, -IQR, -lower_bound, -upper_bound)


interaction_outliers %>% 
  group_by(method, corstr, cov_set) %>% 
  summarise(n = n()) %>% 
  arrange(n) %>% 
  kable(booktabs = TRUE, digits = 3, 
        col.names = c("Method", "Working Correlation", 
                      "Confounders", "No.Outliers"), 
        caption = "Outliers in Relative Bias for the interaction effect between BAV and Time") %>% 
  kable_paper("striped", full_width = F) %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  collapse_rows(columns = 1:2, valign = "top") 
```


## Estimate for $\rho$

QLS method tends to overestimate the correlation between repeated measurements, at around 0.7, while the GEE method produced close estimate when the working correlation is correctly specified. 

```{r}
calculate_mean_rho <- function(df) {
  df %>%
    map_dfr(~ .x %>% select(rho) %>% distinct()) %>%
    summarise(mean_rho = mean(rho), 
              sd_rho = sd(rho)) %>% 
    mutate(bias = mean_rho - 0.3, 
           rel_bias = bias/0.3, 
           mse = sd_rho^2+bias^2)
}
```

```{r}
rho_df <- calculate_mean_rho(gee_sim_results$ar1_mdl_red) %>% 
  rbind(calculate_mean_rho(gee_sim_results$ar1_mdl_full)) %>% 
  rbind(calculate_mean_rho(gee_sim_results$exch_mdl_red)) %>% 
  rbind(calculate_mean_rho(gee_sim_results$exch_mdl_full)) %>% 
  select(mean_rho, bias) %>% 
  cbind(corstr = c(rep("AR1", 2), rep("Exchangeable", 2))) %>% 
  cbind(mod = rep(c("No", "Yes"), 2)) %>% 
  cbind(rho=0.3) %>% 
  relocate(c(corstr, mod, rho), .before = mean_rho) %>% 
  cbind(calculate_mean_rho(qls_sim_results$ar1_mdl_red) %>% 
      rbind(calculate_mean_rho(qls_sim_results$ar1_mdl_full)) %>%
      rbind(calculate_mean_rho(qls_sim_results$exch_mdl_red)) %>% 
      rbind(calculate_mean_rho(qls_sim_results$ar1_mdl_full)) %>% 
      select(mean_rho, bias))

colnames(rho_df) <- c("Working Correlation", "Covariate Set", "Truth", 
                      "Mean Estimate",  "Bias", "Mean Estimate", "Bias")

rho_df %>% 
  kable(digits = 3) %>% 
  kable_paper("striped", full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1, valign = "top") %>% 
  add_header_above(c(" " = 3, "GEE" = 2, "QLS" = 2))
```


### Calculate Tau

```{r}
# Function to extract tau from QLS estimation 
pull_tau <- function(df){
  df %>%
    map_dfr(~ .x %>% select(tau) %>% distinct())  %>% 
    pull(tau)
}
```


```{r}
red_tau <- data.frame(
  ind = pull_tau(qls_sim_results$ind_mdl_red),
  ar1 = pull_tau(qls_sim_results$ar1_mdl_red),
  exch = pull_tau(qls_sim_results$exch_mdl_red)) %>% 
  cbind(covset = "No") 

full_tau <- data.frame(
  ind = pull_tau(qls_sim_results$ind_mdl_full),
  ar1 = pull_tau(qls_sim_results$ar1_mdl_full),
  exch = pull_tau(qls_sim_results$exch_mdl_full)) %>% 
  cbind(covset = "Yes")

tau_df <- red_tau %>% rbind(full_tau) %>% 
  pivot_longer(!covset, names_to = "corstr", values_to = "tau") 

tau_df %>% 
  group_by(corstr, covset) %>% 
  summarise(mean_tau = mean(tau), 
            sd_tau = sd(tau)) %>% 
  mutate(corstr = case_when(corstr == "ar1" ~ "AR1", 
                            corstr == "ind" ~ "Independence", 
                            TRUE ~ "Exchangeable")) %>% 
  kable(digits = 3, escape = F, 
        col.names = c("Working Correlation", "Covariate Set", "Mean $\\tau$", "SD $\\tau$")) %>% 
  kable_paper("striped", full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1, valign = "top") 
```

```{r}
tau_df %>% 
  ggplot(aes(x = corstr, y = tau, color = corstr)) +
  geom_boxplot()+
  facet_grid(~covset) +
  labs(y =expression(tau),
       x = "",
       color = "Working Correlation", 
       caption = "No = no inclusion of confounding covariates,
       Yes = with inclusion confounding covariates") +
  scale_color_brewer(palette = "Pastel1") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"), 
        panel.grid.major.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        strip.placement = "outside", 
        legend.position = "bottom")
```

