---
title: "Simulation Result Analysis"
author: "Jinyu Luo"
date: "2024-04-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, echo=FALSE)
library(tidyverse)
library(kableExtra)
```

```{r}
load(file = "Outputs/raw_res.rds")

# Full model
gee_coef <- c(b0=-3.815, bav=-1.938, visit=0.114, age=-0.036, 
              male=1.196, bsa=1.955, bav_visit=0.345)
param_full <- c("Intercept","BAV","Visit", "Age", "Male", "BSA","BAV:Visit")

# Reduced model 
red_gee_coef <- c(gee_coef[1:3], gee_coef[7])
red_param <- c("Intercept", "BAV", "Visit",  "BAV:Visit")
```


# Coverage Probability 

## Full Model

The coverage probabilities for estimates of the interaction effect all achieved at least 92%. Probabilities from simulated data with dropouts are one percent lower than the results based on the full data, but these probabilities were improved to around 94% (similar to the results of the full data without adjusting for SEs) after adjusting for standard errors. 

```{r}
full_cp <- rbind(
  # Without dropouts 
  colMeans(fullmod_res$independent$Coverage, na.rm = TRUE),
  colMeans(fullmod_res$exchangeable$Coverage,  na.rm = TRUE),
  colMeans(fullmod_res$ar1$Coverage,  na.rm = TRUE), 
  # With dropouts
  colMeans(fullmod_resdo$independent$Coverage, na.rm = TRUE),
  colMeans(fullmod_resdo$exchangeable$Coverage, na.rm = TRUE),
  colMeans(fullmod_resdo$ar1$Coverage, na.rm = TRUE)) 
colnames(full_cp) <- rep(param_full, 2)
rownames(full_cp) <- rep(c("Independent", "Exchangeable", "AR1"), 2) 
```


```{r}
kable(full_cp, escape = FALSE, digits = 3, 
      caption = "True Parameter Coverage Rate from Full Model with 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  add_header_above(c(" ", "Unadjusted Standard Error" = 7, "Adjusted Standard Error" = 7)) %>% 
  column_spec(8, bold = T) %>% 
  column_spec(15, bold = T) %>% 
  pack_rows("Without Drop out", 1, 3) %>% 
  pack_rows("With Drop outs", 4, 6) %>% 
  landscape()
```

\newpage 

## Reduced Model 

There is no significant difference in coverage probabilities between results from the full data and results from data with dropouts. 

```{r}
### 2. Reduced Model ---------------------------------------------
red_cp <- rbind(
  # Without dropouts
  colMeans(redmod_res$independent$Coverage, na.rm = TRUE),
  colMeans(redmod_res$exchangeable$Coverage, na.rm = TRUE),
  colMeans(redmod_res$ar1$Coverage,  na.rm = TRUE), 
  # With dropouts 
  colMeans(redmod_resdo$independent$Coverage, na.rm = TRUE),
  colMeans(redmod_resdo$exchangeable$Coverage, na.rm = TRUE),
  colMeans(redmod_resdo$ar1$Coverage, na.rm = TRUE)) 
colnames(red_cp) <- rep(red_param, 2)
rownames(red_cp) <- rep(c("Independent", "Exchangeable", "AR1"), 2)
```


```{r}
kable(red_cp, escape = FALSE, digits = 3, 
      caption = "True Parameter Coverage Rate from Reduced Model with 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  add_header_above(c(" ", "Unadjusted Standard Error" = 4, "Adjusted Standard Error" = 4)) %>% 
  column_spec(5, bold = T) %>% 
  column_spec(9, bold = T) %>% 
  pack_rows("Without Drop out", 1, 3) %>% 
  pack_rows("With Drop outs", 4, 6)
```


# Check Convergence 

## Visualize the distribution standard errors 

Titles with `_d` followed the correlation structure name indicates that dropout data. 

### Full Model 
```{r}
se_full <- data.frame(
  # Without dropouts
  independence = fullmod_res$independent$SE[,7], 
  exchangeable = fullmod_res$exchangeable$SE[,7], 
  ar1 = fullmod_res$ar1$SE[,7],
  # With dropouts
  independence_d = fullmod_resdo$independent$SE[,7],
  exchangeable_d = fullmod_resdo$exchangeable$SE[,7], 
  ar1_d = fullmod_resdo$ar1$SE[,7])
plots <- lapply(names(se_full), function(col) {
  se_full %>%
    ggplot(aes_string(x = col)) +
    geom_density() +
    ggtitle(paste("Density of", col))
})
do.call(gridExtra::grid.arrange, plots)
```
The density plot suggests that  
1. Non-convergence occurred in exchangeable correlation structure for both full data and with dropouts.  
2. Standard errors are distributed from 0 to 0.5. 

### Reduced Model 
```{r}
se_red <- data.frame(
  # Without dropouts
  independence = redmod_res$independent$SE[,4], 
  exchangeable = redmod_res$exchangeable$SE[,4], 
  ar1 = redmod_res$ar1$SE[,4],
  # With dropouts
  independence_d = redmod_resdo$independent$SE[,4],
  exchangeable_d = redmod_resdo$exchangeable$SE[,4], 
  ar1_d = redmod_resdo$ar1$SE[,4])
plots <- lapply(names(se_red), function(col) {
  se_red %>%
    ggplot(aes_string(x = col)) +
    geom_density() +
    ggtitle(paste("Density of", col))
})
do.call(gridExtra::grid.arrange, plots)
```

The density plot suggests that  
1. Non-convergence occurred in AR1 correlation structure for data with dropouts.  
2. Standard errors are distributed from 0 to 0.5. 

## Convergence Selection 

```{r}
se_max <- 5
bad_full <- se_full %>% mutate_all(funs(. > se_max)) %>% lapply(., which)
bad_full
```

One non-converged simulation was detected in GEE fit with Exchangeable correlation structure from both full data and dropout data. 

```{r}
bad_red <- se_red %>% mutate_all(funs(. > se_max)) %>% lapply(., which)
bad_red
```

One non-converged simulation was detected in GEE fit with AR1 correlation from data with dropouts. 


Remove non-converged simulations and check the standard error distributions. 

```{r, out.width="50%", fig.show='hold', echo=TRUE}
corrstr <- c("Independent", "Exchangeable", "AR1")
for (i in 1:6) {
  # Adjusted for Control Covariates
  if(length(bad_full[[i]]) != 0){
    if(i > 3){
      cor_idx <- i-3
      fullmod_resdo[[cor_idx]] <- lapply(fullmod_resdo[[cor_idx]], 
                                         function(x) x[-bad_full[[i]], ])
      se <- density(fullmod_resdo[[cor_idx]]$SE[,7])
      plot(se,main="Density of Standard Errors Full model with dropouts", 
           sub = corrstr[cor_idx])
      } else{
      fullmod_res[[i]] <- lapply(fullmod_res[[i]], function(x) x[-bad_full[[i]],])
      se <- density(fullmod_res[[i]]$SE[,7])
      plot(se, main="Density of Standard Errors Full model without dropouts", 
           sub = corrstr[i])
    }
  }
  
  # Unadjusted for Control Covariates
  if(length(bad_red[[i]]) != 0){
    if(i > 3){
      cor_idx <- i-3
      redmod_resdo[[cor_idx]] <- lapply(redmod_resdo[[cor_idx]], 
                                         function(x) x[-bad_red[[i]], ])
      se <- density(redmod_resdo[[cor_idx]]$SE[,4])
      plot(se,main="Density of Standard Errors Reduced model with dropouts", 
           sub = corrstr[cor_idx])
      } else{
      redmod_res[[i]] <- lapply(redmod_res[[i]], function(x) x[-bad_full[[i]],])
      se <- density(redmod_res[[i]]$SE[,4])
      plot(se,main="Density of Standard Errors Reduced model without dropouts", 
           sub = corrstr[i])
    }
  }
}
```

The standard error plots suggests that non-converged simulations are all removed. 

\newpage 

# Estimation Results 

## Adjusted for Control Covariates 

### Without Dropouts

```{r}
res_cols <- c("Mean Estimate", "SD(Estimate)", "Mean Std. Error", 
              "SD(Std. Error)", "Mean Bias", "Mean MSE")
full_results <- cbind(`True values` = gee_coef, 
                      # Independent
                      ind.est = colMeans(fullmod_res$independent$Estimation, na.rm = TRUE),
                      ind.sd.est = apply(fullmod_res$independent$Estimation, 2, sd, na.rm = TRUE), 
                      ind.se = colMeans(fullmod_res$independent$SE, na.rm = TRUE),
                      ind.sd.se = apply(fullmod_res$independent$SE, 2, sd, na.rm = TRUE),
                      ind.bias = colMeans(fullmod_res$independent$Bias, na.rm = TRUE),
                      ind.mse = colMeans(fullmod_res$independent$MSE, na.rm = TRUE), 
                      # Exchangeable
                      exch.est = colMeans(fullmod_res$exchangeable$Estimation, na.rm = TRUE),
                      exch.sd.est = apply(fullmod_res$exchangeable$Estimation, 2, sd, na.rm = TRUE),
                      exch.se = colMeans(fullmod_res$exchangeable$SE, na.rm = TRUE),
                      exch.sd.se = apply(fullmod_res$exchangeable$SE, 2, sd, na.rm = TRUE),
                      exch.bias = colMeans(fullmod_res$exchangeable$Bias, na.rm = TRUE),
                      exch.mse = colMeans(fullmod_res$exchangeable$MSE, na.rm = TRUE),
                      # AR1
                      ar1.est = colMeans(fullmod_res$ar1$Estimation, na.rm = TRUE),
                      ar1.sd.est = apply(fullmod_res$ar1$Estimation, 2, sd, na.rm = TRUE),
                      ar1.se = colMeans(fullmod_res$ar1$SE, na.rm = TRUE),
                      ar1.sd.se = apply(fullmod_res$ar1$SE, 2, sd, na.rm = TRUE),
                      ar1.bias = colMeans(fullmod_res$ar1$Bias, na.rm = TRUE),
                      ar1.mse = colMeans(fullmod_res$ar1$MSE, na.rm = TRUE)) %>% t()
colnames(full_results) <- param_full
rownames(full_results) <- c("True Value", rep(res_cols, 3))
kableExtra::kable(full_results, escape = FALSE, digits = 3, 
                  caption = "Estimation Results for Full Model without dropout from 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  column_spec(8, bold = T) %>% 
  pack_rows("Independent", 2, 7) %>% 
  pack_rows("Exchangeable", 8, 13) %>% 
  pack_rows("AR1", 14, 19)
```

**Observations**:  
1. Exchangeable and independent working correlation had the same estimate for the interaction effect.  
2. Estimates for main effects of BAV and Visits are all similar across the three working correlations and similar to true values.  
3. The mean bias and mean MSE are similar, around 0.05, regardless of working correlation structure. 

\newpage 

### With Dropouts

```{r}
full_dropouts <- cbind(
  `True values` = gee_coef, 
  # Independent
  ind.est = colMeans(fullmod_resdo$independent$Estimation, na.rm = TRUE),
  ind.sd.est = apply(fullmod_resdo$independent$Estimation, 2, sd, na.rm = TRUE),
  ind.se = colMeans(fullmod_resdo$independent$SE, na.rm = TRUE),
  ind.sd.se = apply(fullmod_resdo$independent$SE, 2, sd, na.rm = TRUE),
  ind.bias = colMeans(fullmod_resdo$independent$Bias, na.rm = TRUE),
  ind.mse = colMeans(fullmod_resdo$independent$MSE, na.rm = TRUE), 
  # Exchangeable
  exch.est = colMeans(fullmod_resdo$exchangeable$Estimation, na.rm = TRUE),
  exch.sd.est = apply(fullmod_resdo$exchangeable$Estimation, 2, sd, na.rm = TRUE),
  exch.se = colMeans(fullmod_resdo$exchangeable$SE, na.rm = TRUE),
  exch.sd.se = apply(fullmod_resdo$exchangeable$SE, 2, sd, na.rm = TRUE),
  exch.bias = colMeans(fullmod_resdo$exchangeable$Bias, na.rm = TRUE),
  exch.mse = colMeans(fullmod_resdo$exchangeable$MSE, na.rm = TRUE),
  # AR1
  ar1.est = colMeans(fullmod_resdo$ar1$Estimation, na.rm = TRUE),
  ar1.sd.est = apply(fullmod_resdo$ar1$Estimation, 2, sd, na.rm = TRUE),
  ar1.se = colMeans(fullmod_resdo$ar1$SE, na.rm = TRUE),
  ar1.sd.se = apply(fullmod_resdo$ar1$SE, 2, sd, na.rm = TRUE),
  ar1.bias = colMeans(fullmod_resdo$ar1$Bias, na.rm = TRUE),
  ar1.mse = colMeans(fullmod_resdo$ar1$MSE, na.rm = TRUE)) %>% t()
colnames(full_dropouts) <- param_full
rownames(full_dropouts) <- c("True Value", rep(res_cols, 3))
kableExtra::kable(full_dropouts, escape = FALSE, digits = 3, 
                  caption = "Estimation Results for Full Model with Drop outs from 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  column_spec(8, bold = T) %>% 
  pack_rows("Independent", 2, 7) %>% 
  pack_rows("Exchangeable", 8, 13) %>% 
  pack_rows("AR1", 14, 19)
```

**Observations:**  
1. Exchangeable working correlation provided the closest estimates for all coefficients.  
2. Estimates are similar between independent and AR1 working correlations for all coefficients.  
3. Mean biases and mean MSEs are similar, at around 0.03 and 0.120, respectively. 

\newpage 

## Without Adjusting for Control Covariates 

### Without Dropouts

```{r}
# 2. Reduced Covariate set 
reduced_result <- cbind(
  `True values` = red_gee_coef, 
  # Independent
  ind.est = colMeans(redmod_res$independent$Estimation, na.rm = TRUE),
  ind.sd.est = apply(redmod_res$independent$Estimation, 2, sd, na.rm = TRUE),
  ind.se = colMeans(redmod_res$independent$SE, na.rm = TRUE),
  ind.sd.se = apply(redmod_res$independent$SE, 2, sd, na.rm = TRUE),
  ind.bias = colMeans(redmod_res$independent$Bias, na.rm = TRUE),
  ind.mse = colMeans(redmod_res$independent$MSE, na.rm = TRUE), 
  # Exchangeable
  exch.est = colMeans(redmod_res$exchangeable$Estimation, na.rm = TRUE),
  exch.sd.est = apply(redmod_res$exchangeable$Estimation, 2, sd, na.rm = TRUE),
  exch.se = colMeans(redmod_res$exchangeable$SE, na.rm = TRUE),
  exch.sd.se = apply(redmod_res$exchangeable$SE, 2, sd, na.rm = TRUE),
  exch.bias = colMeans(redmod_res$exchangeable$Bias, na.rm = TRUE),
  exch.mse = colMeans(redmod_res$exchangeable$MSE, na.rm = TRUE),
  # AR1
  ar1.est = colMeans(redmod_res$ar1$Estimation, na.rm = TRUE),
  ar1.sd.est = apply(redmod_res$ar1$Estimation, 2, sd, na.rm = TRUE),
  ar1.se = colMeans(redmod_res$ar1$SE, na.rm = TRUE),
  ar1.sd.se = apply(redmod_res$ar1$SE, 2, sd, na.rm = TRUE),
  ar1.bias = colMeans(redmod_res$ar1$Bias, na.rm = TRUE),
  ar1.mse = colMeans(redmod_res$ar1$MSE, na.rm = TRUE)) %>% t()
colnames(reduced_result) <- red_param
rownames(reduced_result) <- c("True Value", rep(res_cols, 3))
kableExtra::kable(reduced_result, escape = FALSE, digits = 3, 
                  caption = "Estimate Results for Reduced Model without dropouts from 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  column_spec(5, bold = T) %>% 
  pack_rows("Independent", 2, 7) %>% 
  pack_rows("Exchangeable", 8, 13) %>% 
  pack_rows("AR1", 14, 19)
```

**Observations:**  
The three working correlations provided the exact same estimate for the interaction effect and the main effect of Visit, but the independent working correlation had slightly larger mean bias, at -0.025.  

\newpage 

### With Dropouts

```{r}
# Reduced model with drop outs 
reduced_dropouts <- cbind(
  `True values` = red_gee_coef, 
  # Independent
  ind.est = colMeans(redmod_resdo$independent$Estimation, na.rm = TRUE),
  ind.sd.est = apply(redmod_resdo$independent$Estimation, 2, sd, na.rm = TRUE),
  ind.se = colMeans(redmod_resdo$independent$SE, na.rm = TRUE),
  ind.sd.se = apply(redmod_resdo$independent$SE, 2, sd, na.rm = TRUE),
  ind.bias = colMeans(redmod_resdo$independent$Bias, na.rm = TRUE),
  ind.mse = colMeans(redmod_resdo$independent$MSE, na.rm = TRUE), 
  # Exchangeable
  exch.est = colMeans(redmod_resdo$exchangeable$Estimation, na.rm = TRUE),
  exch.sd.est = apply(redmod_resdo$exchangeable$Estimation, 2, sd, na.rm = TRUE),
  exch.se = colMeans(redmod_resdo$exchangeable$SE, na.rm = TRUE),
  exch.sd.se = apply(redmod_resdo$exchangeable$SE, 2, sd, na.rm = TRUE),
  exch.bias = colMeans(redmod_resdo$exchangeable$Bias, na.rm = TRUE),
  exch.mse = colMeans(redmod_resdo$exchangeable$MSE, na.rm = TRUE),
  # AR1
  ar1.est = colMeans(redmod_resdo$ar1$Estimation, na.rm = TRUE),
  ar1.sd.est = apply(redmod_resdo$ar1$Estimation, 2, sd, na.rm = TRUE),
  ar1.se = colMeans(redmod_resdo$ar1$SE, na.rm = TRUE),
  ar1.sd.se = apply(redmod_resdo$ar1$SE, 2, sd, na.rm = TRUE),
  ar1.bias = colMeans(redmod_resdo$ar1$Bias, na.rm = TRUE),
  ar1.mse = colMeans(redmod_resdo$ar1$MSE, na.rm = TRUE)) %>% t()
colnames(reduced_dropouts) <- red_param
rownames(reduced_dropouts) <- c("True Value", rep(res_cols, 3))
kableExtra::kable(reduced_dropouts, escape = FALSE, digits = 3, 
                  caption = "Estimate Results for Reduced Model with dropout from 1000 Simulations") %>%
  kable_styling(full_width = F, position = "center") %>% 
  column_spec(5, bold = T) %>% 
  pack_rows("Independent", 2, 7) %>% 
  pack_rows("Exchangeable", 8, 13) %>% 
  pack_rows("AR1", 14, 19)
```

**Observations**:  
Three working correlations provided similar estimates. 
